---
title: "Annotated Eye and Retina"
output: html_notebook
---

## Integration script

integration_from_eye_and_retina_remove_red_blood_cells_increaseMT.Rmd


## Filtering applied : 

Any cells which expressed haemoglobin genes were removed from the analysis. 

The mitochondrial threshold 10%

Any samples with more than 4000 cells downsampled to 4000.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, libraries}

library(Seurat)
library(ggplot2)
library(cowplot)
library(SCFunctionsV3)
library(dplyr)
library(tidyr)
library(stringr)
library(tidyverse)
library(DoubletFinder)
library(pheatmap)
library(harmony)


### Functions
howManyCells <- function(sObj){dim(sObj)[2]}



#sampleInfo <- readRDS("../fetal/rObjects/sampleInfo.rds")


```


## Manual removal clusters

After integration the following clusters were removed:

36- dominated by 16 PCW not present in other stages
40-mix of markers for lens, Acs and fibroblasts
41-dominated by 16 PCW
42- dominated by 16 pcw (lens markers)

```{r, eval = FALSE}
seuratObj <- readRDS("rObjects/seuratObj_retina_eye_mt10_noHB.rds")
seuratObj_manual_filtered <- seuratObj[,!seuratObj@active.ident %in% c(36,40,41,42)]
```



The remaining clusters were annotated as follows:





```{r, eval = FALSE}
annotations <- read.csv("csvFiles/retina_eye_annotations.txt", sep = "-", header = FALSE)
colnames(annotations) <- c("Cluster", "Annotation")



seuratObj_manual_filtered$annotation <- annotations$Annotation[match(seuratObj_manual_filtered$RNA_snn_res.2.2, annotations$Cluster)]


seuratObj_annotated <- seuratObj_manual_filtered
saveRDS(seuratObj_annotated, "rObjects/seuratObj_retina_eye_annotated.rds")

```


```{r}
DimPlot(seuratObj_annotated, 
        group.by = "annotation", 
        label = TRUE) + ggsave("images/annotated_data.png")
```


