---
title: "Whole Eye and Retina  samples"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, libraries}

library(Seurat)
library(cowplot)
library(SCFunctionsV3)
library(dplyr)
library(tidyr)
library(stringr)
library(tidyverse)
library(DoubletFinder)
library(pheatmap)
library(harmony)


### Functions
howManyCells <- function(sObj){dim(sObj)[2]}



#sampleInfo <- readRDS("../fetal/rObjects/sampleInfo.rds")

seuratObj <- readRDS("rObjects/seuratObj_retina_eye_mt10_noHB.rds")
```


## Filtering of Red blood cells and mitochondrial reads

Any cells which expressed haemoglobin genes were removed from the analysis. The mitochondrial threshold was decreased from 20% to 10%

Any samples with more than 4000 cells downsampled to 4000.

```{r}
#sampleInfo 
```



```{r, D10, eval = FALSE}
## filter_hb_mt_integrate_find_markers_Retina_eye.R
### Functions
howManyCells <- function(sObj){dim(sObj)[2]}
## downsample to 4000
downsample <- function(sObj, no_cells = 4000){
  if (howManyCells(sObj) >  no_cells){
    set.seed(1234)
    s <- sample(1:howManyCells(sObj), no_cells, replace = FALSE)
    sObj<- sObj[,s]}
  return(sObj)
}

## read in QCed data
seuratObj_retina <- readRDS("../fetal/rObjects/sObj_Filtered.rds")
seuratObj_eye <- readRDS("../../Eye/rObjects/sObj_eye.rds")
seuratObj_list <- c(seuratObj_retina, seuratObj_eye)


## remove, HB cells, change mitchondrial threshold to 10% and downsample to 4000
sObj_filtered <- list()
## remove hb and set mitochondrial threshold to 10%
dontuse <- c("HBP1", "HBEGF", "HBS1L")



for (i in 1:length(seuratObj_list)){
  
  sObj <- seuratObj_list[[i]]
  HB_sample <- rownames(seuratObj_list[[i]])[grepl("^HB", rownames(seuratObj_list[[i]]))]
  HB_sample <- setdiff(HB_sample, dontuse)
  
  if (length(HB_sample) == 1){
    numberHB <- sObj@assays$RNA@counts[HB_sample, ]
  } else{
    numberHB <- sObj@assays$RNA@counts[HB_sample, ] %>% colSums()
  }
  
  ## remove high mitochondrial
  filter <- sObj$percent.mt < 10 & 
    numberHB == 0
  
  sObj <- sObj[,filter]
  
  
  # downsample to 4000
  sObj_filtered[[i]] <- downsample(sObj)
}


## merge objects
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
sObj_ds <- merge(sObj_filtered[[1]], sObj_filtered[c(1:16, 18:24)]) %>% 
  CellCycleScoring(s.features = s.genes, 
                   g2m.features = g2m.genes, 
                   set.ident = FALSE)

sObj_ds$version <- 3
sObj_ds$version[grepl(14149, sObj_ds$sample)]


seuratObj <- sObj_ds %>%
  Seurat::NormalizeData(verbose = FALSE) %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
  ScaleData(verbose = FALSE, 
            vars.to.regress = c("percent.mt", 
                                "nCount_RNA", 
                                "nFeature_RNA")) %>% 
  RunPCA(pc.genes = seuratObj_ds@var.genes, npcs = 20, verbose = FALSE) %>%
  RunHarmony(c("sample", "version"), plot_convergence = TRUE) %>% 
  RunUMAP(reduction = "harmony", dims = 1:10) %>% 
  FindNeighbors(reduction = "harmony", dims = 1:10) %>% 
  FindClusters(resolution = c(0.2,1,2.2))


#seuratObj$stage.2 <- seuratObj$stage
seuratObj$stage <- seuratObj$stage %>% gsub("PCW.*", "PCW", .) 
seuratObj$tissue <- ifelse(is.na(seuratObj$sample), yes = "eye", no = "retina")
seuratObj$sample[is.na(seuratObj$sample)] <- seuratObj$orig.ident[is.na(seuratObj$sample)]

saveRDS(seuratObj, "rObjects/seuratObj_retina_eye_mt10_noHB.rds")

```

```{r, eval = FALSE}

## Don't use
library(clustree)

seuratObj[["RNA_snn_res.0.4"]] <- NULL
seuratObj[["RNA_snn_res.0.6"]] <- NULL
seuratObj[["RNA_snn_res.0.8"]] <- NULL
seuratObj[["RNA_snn_res.1"]] <- NULL

seuratObj[["RNA_snn_res.1.4"]] <- NULL
seuratObj[["RNA_snn_res.1.6"]] <- NULL
seuratObj[["RNA_snn_res.1.8"]] <- NULL
seuratObj[["RNA_snn_res.2"]] <- NULL



p_clus <- clustree(seuratObj,prefix = "RNA_snn_res." )

p_clus_res2.2 <- p_clus$data %>% filter(RNA_snn_res. == 2.2)

```

## Check for technical differences

There was variable proportions of mitochondrial reads and number of genes and counts across samples. These features were regressed out and harmony batch correction applied.

```{r}
seuratObj$sample <- paste(seuratObj$sample, seuratObj$tissue, sep = "_")
VlnPlot(seuratObj, group.by = "sample", features = c("nCount_RNA", "nFeature_RNA", "percent.mt"), pt.size = 0)
```





## Integrated data

There is a lot of overlap between the datasets. 

```{r, fig.width=25, fig.height = 15}
p1 <- DimPlot(seuratObj, group.by = "sample")
p2 <- DimPlot(seuratObj, label = TRUE)
p3 <- DimPlot(seuratObj, group.by = "Phase")
p4 <- DimPlot(seuratObj, group.by = "stage")
p2 + p3 + p1 + p4 

```



```{r}
number_cells <- seuratObj[[]] %>% select(c("RNA_snn_res.2.2")) %>% group_by(RNA_snn_res.2.2) %>% summarise(number_cells = n()) %>%
  ggplot(aes(RNA_snn_res.2.2, number_cells, fill = RNA_snn_res.2.2)) + 
  geom_bar(stat =  "identity")


```



## Numbers of Cells for each stage

The cell numbers were normalised for each developmental stage and then the percentage of cells from each stage within a cluster was calulated. There are a number of clusters dominated by 16PCW.

```{r, fig.height = 15, fig.width= 15}
number_cells <- seuratObj[[]] %>% select(c("stage", "RNA_snn_res.2.2")) %>% group_by(stage, RNA_snn_res.2.2) %>% summarise(number_cells = n()) %>% dplyr::rename(cluster = "RNA_snn_res.2.2")

numberStage <- seuratObj[[]] %>% select(c("stage")) %>% group_by(stage) %>% summarise(number_per_stage = n())

number_cells <- number_cells %>% left_join(numberStage)

number_cells$percentage_cells <- number_cells$number_cells/number_cells$number_per_stage * 100

number_cells <- number_cells %>% group_by(cluster) %>% summarise(normalised_number_cells_per_cluster= sum(percentage_cells)) %>%  right_join(number_cells)

number_cells$normalised_number_cells <- number_cells$percentage_cells / number_cells$normalised_number_cells_per_cluster * 100



number_cells %>% ggplot(aes(stage, normalised_number_cells, fill = stage)) + 
  geom_bar(stat =  "identity") +
  facet_wrap(~cluster, scales="free_y", drop = FALSE) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  



```

## Numbers of Cells for each sample

```{r, fig.height = 15, fig.width= 15}
number_cells <- seuratObj[[]] %>% select(c("sample", "RNA_snn_res.2.2")) %>% group_by(sample, RNA_snn_res.2.2) %>% summarise(number_cells = n()) %>% dplyr::rename(cluster = "RNA_snn_res.2.2")

numbersample <- seuratObj[[]] %>% select(c("sample")) %>% group_by(sample) %>% summarise(number_per_sample = n())

number_cells <- number_cells %>% left_join(numbersample)

number_cells$percentage_cells <- number_cells$number_cells/number_cells$number_per_sample * 100

number_cells <- number_cells %>% group_by(cluster) %>% summarise(normalised_number_cells_per_cluster= sum(percentage_cells)) %>%  right_join(number_cells)

number_cells$normalised_number_cells <- number_cells$percentage_cells / number_cells$normalised_number_cells_per_cluster * 100



number_cells %>% ggplot(aes(sample, normalised_number_cells, fill = sample)) + 
  geom_bar(stat =  "identity") +
  facet_wrap(~cluster, scales="free_y", drop = FALSE) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  



```




```{r, eval = FALSE}
markers <- FindAllMarkers(seuratObj, logfc.threshold = 0.7)## DE

markers <- markers %>% dplyr::arrange(cluster, desc(avg_log2FC))
  
fName <- paste0("csvFiles/", "retina_noHB", ".csv")
write.csv(markers, fName)
fName <- gsub("csvFiles", "rObjects", fName) %>% gsub(".csv", ".rds", .)
saveRDS(markers, fName)

```


```{r}
GeneUMAP <- function(gene, seuratObj, celltype){
  p1 <- FeaturePlot(seuratObj, features = gene, 
                    min.cutoff = "q1", 
                    cols = c("#E5E5E5", "#FF0000"), 
                    label = TRUE, order = TRUE, label.size = 2
                    ) +
    theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        text = element_text(size=10)) 
    
 
# p1 <- p1 + ggsave(paste0("highRes/umap_gene_plots/", celltype, gene, ".tiff"), width = 3, height = 3)
return(p1)
}


RPC_EyeField <- c("LIN28A", "SFRP2", "SFRP1", "SOX2","MKI67", "HES5", "HMGA1", "HMGA2", "HMGB1", "HMGB2", "FZD5", "DLK1", "SPP1", "CDK1", "PAX6", "RAX")
RPC_EyeField <- intersect(RPC_EyeField, rownames(seuratObj))
RPC_list <- lapply(RPC_EyeField, GeneUMAP, seuratObj = seuratObj, celltype = "RPC_EyeField_")


RGCs <- c("GAP43", "NEUROD6", "STMN2", "SNCG", "ELAVL3", "POU4F2")
#setdiff(RGCs, rownames(seuratObj))
RGCs_list <- lapply(RGCs, GeneUMAP, seuratObj = seuratObj, celltype = "RGCs_")


MG <- c("CLU", "VIM", "RLBP1", "CRYAB", "CRABP1")
#setdiff(MG, rownames(seuratObj))
MG_list <- lapply(MG, GeneUMAP, seuratObj = seuratObj, celltype = "MG_")


HCs <- c("ONECUT1", "ONECUT2", "PROX1")
#setdiff(HCs, rownames(seuratObj))
HCs_list <- lapply(HCs, GeneUMAP, seuratObj = seuratObj, celltype = "HCs_")


ACs <- c("TFAP2A", "LHX9")
#setdiff(ACs, rownames(seuratObj))
ACs_list <- lapply(ACs, GeneUMAP, seuratObj = seuratObj, celltype = "ACs_")

# cluster0 <- c("PAX6", "SOX2", "RAX", "VSX2", "OTX1", "AQP1", "ZIC1", "FUT4", "RDH10", "CDO1", "MSX1", "WNT2B", "SFRP1", "SFRP2", "FGF3", "ETV1", "ETV6", "FOXP1", "LIN28A")
# 
# #setdiff(cluster0, rownames(seuratObj))
# cluster0_list <- lapply(cluster0, GeneUMAP, seuratObj = seuratObj, celltype = "cluster0_")

Forebrain_NSC <- c("EMX2", "FGFR3", "FOXG1", "FOXM1")
#setdiff(Forebrain_NSC, rownames(seuratObj))
Forebrain_NSC_list <- lapply(Forebrain_NSC, GeneUMAP, seuratObj = seuratObj, celltype = "Forebrain_NSC_markers_")

Astrocytes <- c("ALDH1A1", "GFAP", "ANGPT1")
#setdiff(Astrocytes, rownames(seuratObj))
Astrocytes_list <- lapply(Astrocytes, GeneUMAP, seuratObj = seuratObj, celltype = "Astrocytes_")

RET_neuroepi <- c("LHX5", "NNAT")
#setdiff(RET_neuroepi, rownames(seuratObj))
RET_neuroepi_list <- lapply(RET_neuroepi, GeneUMAP, seuratObj = seuratObj, celltype = "RET_neuroepi_")


PSCs <- c("TDGF1", "POU5F1")
#setdiff(PSCs, rownames(seuratObj))
PSCs_list <- lapply(PSCs, GeneUMAP, seuratObj = seuratObj, celltype = "PSCs_")

RODS <- c("NRL", "GNAT1", "PDE6G", "NR2E3", "RHO", "GNGT1")
RODS <- intersect(RODS, rownames(seuratObj))
RODS_list <- lapply(RODS, GeneUMAP, seuratObj = seuratObj, celltype = "RODS_")


RPE <- c("TYR", "BEST1", "PMEL", "RPE65", "TYRP1", "MITF")
#setdiff(RPE, rownames(seuratObj))
RPE_list <- lapply(RPE, GeneUMAP, seuratObj = seuratObj, celltype = "RPE_")


CONES <- c("GNB3", "RXRG", "ARR3", "CRX", "RCVRN", "PDE6H")
#setdiff(CONES, rownames(seuratObj))
CONES_list <- lapply(CONES, GeneUMAP, seuratObj = seuratObj, celltype = "CONES_")


T3 <- c("FABP7", "OTX2", "DLL3")
#setdiff(T3, rownames(seuratObj))
T3_list <- lapply(T3, GeneUMAP, seuratObj = seuratObj, celltype = "T3_")

 
T1 <- c("ATOH7", "HES6", "DLL3")
#setdiff(T1, rownames(seuratObj))
T1_list <- lapply(T1, GeneUMAP, seuratObj = seuratObj, celltype = "T1_")



T2 <- c("PRDM13", "PTF1A", "HSPB1")
#setdiff(T2, rownames(seuratObj))
T2_list <- lapply(T2, GeneUMAP, seuratObj = seuratObj, celltype = "T2_")

OSE <- c("KRT8", "KRT18", "CXCL14", "KRT15", "KRT19", "KRT7", "KRT13")
OSE <- intersect(OSE, rownames(seuratObj))
OSE_list <- lapply(OSE, GeneUMAP, seuratObj = seuratObj, celltype = "OSE_")

BP <- c("VSX1", "VSX2", "CADPS", "GRM6", "GRIK1")
#setdiff(BP, rownames(seuratObj))
BP_list <- lapply(BP, GeneUMAP, seuratObj = seuratObj, celltype = "BP_")




```

## FGFR3 and RAX 
```{r}
GeneUMAP("RAX", seuratObj)
GeneUMAP("FGFR3", seuratObj)
```


## RPCs

```{r}
RPC_list

```

## RGCs

```{r}
RGCs_list
```

## MG

```{r}
MG_list
```

## HCs

```{r}
HCs_list
```

## ACs

```{r}
ACs_list
```

## Forebrain

```{r}
Forebrain_NSC_list
```


## Astrocytes

```{r}
Astrocytes_list
```

## PSCs

```{r}
PSCs_list
```

## Rods

```{r}
RODS_list
```

## RPE

```{r}
RPE_list
```

## Cones

```{r}
CONES_list
```

## T3

```{r}
T3_list
```

## T1

```{r}
T1_list
```

## T2

```{r}
T2_list
```

## OSE

```{r}
OSE_list
```

## BP

```{r}
BP_list
```

