---
title: "`r thetitle`"
output:
  html_document:
    df_print: paged
---

```{r}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r libaries}
library(dplyr)
library(Seurat)
library("ggplot2")
library(monocle3)
library(patchwork)
library(SeuratWrappers)
library(harmony)
library(ggplotify)
sobjToCds <- function(sObj){
  harmonyLoadings <- sObj@reductions[["harmony"]]@feature.loadings.projected
  sObj <- DietSeurat(sObj, graphs = "umap")
  cds <- as.cell_data_set(sObj)
  cds <- cluster_cells(cds, reduction_method = "UMAP", k = 15)
  cds <- learn_graph(cds, use_partition = FALSE)
  ## Add gene names into CDS
  rowData(cds)$gene_name <- rownames(cds)
  rowData(cds)$gene_short_name <- rowData(cds)$gene_name
  ## estimate size factors
  cds <- estimate_size_factors(
  cds,
  round_exprs = TRUE,
  method = c("mean-geometric-mean-total", "mean-geometric-mean-log-total")
)
  # add harmony reduction
  cds@preprocess_aux$gene_loadings <- harmonyLoadings
  
  
return(cds)
}

seuratObj <- readRDS("../rObjects/seuratObj_retina_eye_annotated.rds")
## add consensus clusters
cellAnnoWide <- readRDS("rObjects/cellAnnoWide.rds")
cellAnnoWide <- cellAnnoWide[colnames(seuratObj), ]
cellAnnoWide$consensus[seuratObj@active.ident == 30] <- "RGCs"
seuratObj$consensus <- cellAnnoWide$consensus %>% gsub("early ", "", .) %>% gsub("late ", "", .) 
seuratObj$consensus <- seuratObj$consensus %>% gsub("removed_", "", .) 

seuratObj_subset <- readRDS(paste0("rObjects/seurat_subset_", outName, ".rds"))
cds_subset_cc <- readRDS(paste0("rObjects/cds_", outName, ".rds"))

subsets <- readRDS("rObjects/subsets_list_consensus.rds")
startingNode <- read.csv("StartNode/startingNodes", row.names = 1)[outName,]

```



# With cell cycle regression


```{r}
clustersUse <- subsets[[outName]]

## subset_trajectory.R
```

Cells from  `r paste0(clustersUse, collapse = ",")` were selected. The cells selected are shown below:

```{r}
seuratObj$cellSelected <- NA
seuratObj$cellSelected[colnames(seuratObj_subset)] <- seuratObj_subset$consensus
DimPlot(seuratObj, group.by = "cellSelected")
```


### Clustering and annotation

Cell cycle regression and harmony batch correction performed.The new clusters are shown below:  




```{r, fig.width=20, fig.height = 15}
p1 <- DimPlot(seuratObj_subset, label = TRUE, group.by = "consensus") +   ggtitle("new annotation") 

p2 <- DimPlot(seuratObj_subset, group.by = "RNA_snn_res.1", label = TRUE)  + ggtitle("new clusters")

p1 + p2 
```

## Pseudotime analysis

Monocle 3 was used to learn the trajectory in the data and the cells were ordered by pseudotime. Node `r startingNode`, was used as the starting point for the trajectory. 

```{r}
# convert_seurat_to_cds.R 
```

```{r, fig.width = 10, fig.height=10}
cds_subset_cc <- order_cells(cds_subset_cc, root_pr_nodes = startingNode)
plot_cells(cds_subset_cc, 
           label_principal_points = TRUE, 
           color_cells_by = "annotation"
           )
```


## Pseudotime Ordering

The pseudotime ordering is shown below.

```{r, fig.width = 10, fig.height=10}
plot_cells(cds_subset_cc, 
           label_principal_points = TRUE, 
           color_cells_by = "pseudotime"
           )
```


## Markers for each cluster

Markers were then found for each of the new clusters. 

```{r, eval = FALSE}
# find_markers.R
```

```{r, fig.width= 20,fig.height=20}

seuratObj_subset <- SetIdent(seuratObj_subset, value = "consensus")
cellTypes <- seuratObj_subset$consensus %>% unique()

cellTypeOrder <- read.csv("annotations/cellTypeOrder.csv", 
                          header = FALSE, 
                          stringsAsFactors = FALSE)$V1 

clusterOrder <- cellTypeOrder[cellTypeOrder %in% cellTypes]


library(pheatmap)
library(viridis)


markers <- read.csv(paste0("csvFiles/", outName, ".csv"), row.names = 1)
markers$cluster <- markers$cluster %>%  factor(levels = clusterOrder )

seuratObj_subset$consensus2 <- seuratObj_subset$consensus %>% factor(levels = clusterOrder)


seuratObj_subset$pseudotime <- pseudotime(cds_subset_cc)
cellOrder <- order(seuratObj_subset$consensus2, rev(seuratObj_subset$pseudotime))

top10 <- markers %>% 
  filter(avg_log2FC > 0) %>% 
  arrange(cluster) %>% 
  group_by(cluster) %>% 
  top_n(n = 10, wt = avg_log2FC) 

#top10 <- top10[c(-1,-2,-3),]

genesUse <- top10 %>% pull(gene) %>% unique()




md <- seuratObj_subset[[]][cellOrder,c("consensus", "pseudotime", "stage")]




md$stage <-md$stage %>% gsub("PCW", "", .) %>% as.numeric()

breaksList = seq(0, 3, by = .2)


exp <- seuratObj_subset@assays$RNA@data[genesUse, cellOrder]

p_heatmap1 <- as.ggplot(pheatmap(exp, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         show_colnames = FALSE, 
         annotation_col = md, 
         color = viridis(15), 
         scale = "row", 
         breaks = breaksList
          ))




```

