---
title: "Monocle ATAC all"
output:
  html_document:
    df_print: paged
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```


```{r, libraries}
library(tidyverse)
library(Signac)
library(Seurat)
library(JASPAR2020)
library(GenomeInfoDb)
library(EnsDb.Hsapiens.v86)
library(TFBSTools)
library(monocle3)
library(SeuratWrappers)
library(patchwork)

pfm <- getMatrixSet(
  x = JASPAR2020,
  opts = list(collection = "CORE", tax_group = 'vertebrates', all_versions = FALSE)
)

# motifs <- sapply(pfm, function(x){x@name})
# motif_df <- data.frame(motif = names(motifs),
#            name = unname(motifs)
#            )
# 
# 
# retina_markers <- readRDS("../markers/Retina_Markers.rds")
# #retina_markers_df <-
# 
# 
# motif_df$motif[match(retina_markers[[2]], motif_df$name)]
motifs <- readRDS("rObjects/shared_enriched_motifs_with13PCW_cellTypes.rds")
sObj <- readRDS("rObjects/shared_sObj_merge_with13PCW_footprints.rds")
# DimPlot(integrated, label = TRUE)


motif_expr <- sObj@assays$chromvar@data
motif_expr <- motif_expr[rowSums(motif_expr) > 0 ,]
sObj$nMotif <- motif_expr %>% colSums()
```


## Annotate cells

The cells were annotated using the annotations on slide 155 with new AC definitions:

```{r}
anno <- read.csv("annotation/all_samples")

sObj$CellType <- anno$cell_type[match(sObj$peaks_snn_res.0.8, anno$cluster)]
DimPlot(sObj, group.by = "CellType", label = TRUE)

FeaturePlot(sObj, "nMotif", label = TRUE, max.cutoff = 2000, min.cutoff = 0)
FeaturePlot(sObj, "MA0148.4", min.cutoff = 0)

```



## RPC T1 T2 T2 trajectory
```{r}
start_Node = "Y_31"
```

The RPC, T1, T2, T3 cell types were selected and Monocle 3 pseudotime analysis performe starting from `r start_Node`.  

```{r, eval = FALSE}

RPC_T1_T2_T3 <- sObj[,  sObj$CellType %in% c("early RPCs", "late RPCs", "T1", "T2", "T3")]


RPC_T1_T2_T3[,RPC_T1_T2_T3$nCount_peaks > 25000]


RPC_T1_T2_T3_new <- RunUMAP(object = RPC_T1_T2_T3,
                      reduction = 'lsi',
                      dims = 2:10,
                      n.neighbors = 30,
                      min.dist = 0.25,
                      n.epochs = 5000
                      )



saveRDS(RPC_T1_T2_T3_new, "rObjects/seurat_PT_RPC_T1_T2_T3.rds")



```

```{r}

RPC_T1_T2_T3_new <- readRDS("rObjects/seurat_PT_RPC_T1_T2_T3.rds")

cds <- as.cell_data_set(RPC_T1_T2_T3_new)
cds <- cluster_cells(cds = cds, reduction_method = "UMAP")
cds <- learn_graph(cds, use_partition = FALSE, close_loop = FALSE)
cds <- order_cells(cds, reduction_method = "UMAP", root_pr_nodes = start_Node)

### add pseudotime back to seurat object
sObj <- AddMetaData(
  object = sObj,
  metadata = cds@principal_graph_aux@listData$UMAP$pseudotime,
  col.name = "RPC_T1_T2_T3"
)



```

```{r}
p1 <- plot_cells(
  cds = cds,
  show_trajectory_graph = TRUE,
  color_cells_by = "CellType", label_principal_points = TRUE, cell_stroke = .5
) 





p0 <- p1
p0$layers[[1]] <- NULL
p0
```



```{r}
p0 <-  plot_cells(
  cds = cds,
  show_trajectory_graph = TRUE,
  color_cells_by = "pseudotime", label_principal_points = TRUE,  cell_stroke = .5
) 

p0$layers[[1]] <- NULL
p0

```

The pseudotime estimated are shown on the original umap below:


```{r}
DimPlot(sObj, group.by = "CellType", label = TRUE)

FeaturePlot(sObj, c("RPC_T1_T2_T3"), pt.size = 0.1) & scale_color_viridis_c()

```



The correlation between pseudotime between motifs enrichment was calculated:

```{r}
cellTypes <- names(motifs)
pseudotime <-sObj$RPC_T1_T2_T3[!is.na(sObj$RPC_T1_T2_T3)]
cor_row <- function(x){cor(pseudotime, x) }


addCorrelation <- function(cellType){
motif_mat <- sObj@assays$chromvar@data[motifs[[cellType]]$motif,!is.na(sObj$RPC_T1_T2_T3)]
motifs[[cellType]]$corr_RPC_T1_T2_T3 <- apply(motif_mat, 1, cor_row)
return(motifs[[cellType]])
}


motifs <- lapply(names(motifs), addCorrelation)
names(motifs) <- cellTypes

```


## Top Motif plots {.tabset .tabset-pills}

The motif lists were filtered to include only motifs in at least 30% of the cells that there were enriched for with a log fold change of over 2. The top 5 absolute correlation scores for each cell type then selected


```{r}
motifs_filtered <- lapply(motifs, function(x) {x[x$fold.enrichment > 2 & x$percent.observed > 30, ]})

motif_RPC <- motifs_filtered[["early RPCs"]]

findTopMotifs <- function(x){x %>% arrange(desc(abs(corr_RPC_T1_T2_T3))) %>% head(n = 5)}

topMotifs <- lapply(motifs_filtered, findTopMotifs)

# 
# 
# no_motifs <- motifs_filtered %>% sapply(nrow) %>% data.frame() 
# colnames(no_motifs) <- "number of motifs"
# 
# no_motifs

```

```{r}
MotifFeaturePlot <- function(motif){
  motif_name <- motifs[[1]]$motif.name[motifs[[1]]$motif == motif]
  p1 <- FeaturePlot(sObj, motif, min.cutoff = 0, cols = c("lightgrey", "red"), order = TRUE, pt.size = 1 ) + ggtitle( paste(motif_name, motif))
  p2 <- MotifPlot(sObj, motif )
 layout <- '
A#
AB
'
p3 <- p1 + p2   +
  plot_layout(design = layout, widths = c(2,1))
  return(p3)
}



DefaultAssay(sObj) <- "peaks"
```

```{r}
footprints <- GetAssayData(object = sObj, assay = "peaks", 
              slot = "positionEnrichment") %>% names()


p2 <- PlotFootprint(sObj, features = c(topMotifs[["early RPCs"]]$motif.name))
```


### early RPCs

```{r, fig.width=20, fig.height=10}
lapply(topMotifs[["early RPCs"]]$motif, MotifFeaturePlot)
```

```{r}
T1 <- topMotifs[["T1"]]$motif
T2 <- topMotifs[["T2"]]$motif
T3 <- topMotifs[["T3"]]$motif



T1_T2 <- intersect(T1, T2)
T1_T3 <- intersect(T1, T3)

T1_T2_T3 <- intersect(T3, T1_T2)
T1_T2 <- setdiff(T1_T2, T1_T2_T3)
T1_T3 <- setdiff(T1_T3, T1_T2_T3)

T1_only <- setdiff(T1, c(T2,T3))
T2_only <- setdiff(T2, c(T1,T3))
T3_only <- setdiff(T3, c(T2,T1))

T1 <- intersect(T1, T1_T2)


# T1_T2_T3 <- c(topMotifs[["T1"]]$motif.name, topMotifs[["T2"]]$motif.name, topMotifs[["T3"]]$motif) %>% unique()




```

### T1 only

No motifs were found that were unique to T1


### T1 and T2


```{r, fig.width=20, fig.height=10}
lapply(T1_T2, MotifFeaturePlot)
```

### T2 only

```{r, fig.width=20, fig.height=10}
lapply(T2_only, MotifFeaturePlot)
```

### T1, T2 and T3 

```{r, fig.width=20, fig.height=10}
lapply(T1_T2_T3, MotifFeaturePlot)
```


### T1 and T3

```{r, fig.width=20}
lapply(T1_T3, MotifFeaturePlot)
```


### T3 only 


```{r, fig.width=20, fig.height=10}
lapply(T3, MotifFeaturePlot)
```

