---
title: "Monocle ATAC all"
output:
  html_document:
    df_print: paged
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```


```{r, libraries}
library(tidyverse)
library(Signac)
library(Seurat)
library(JASPAR2020)
library(GenomeInfoDb)
library(EnsDb.Hsapiens.v86)
library(TFBSTools)
library(monocle3)
library(SeuratWrappers)
library(patchwork)

pfm <- getMatrixSet(
  x = JASPAR2020,
  opts = list(collection = "CORE", tax_group = 'vertebrates', all_versions = FALSE)
)

motifs <- sapply(pfm, function(x){x@name})
motif_df <- data.frame(motif = names(motifs),
           name = unname(motifs)
           )


retina_markers <- readRDS("../markers/Retina_Markers.rds")
#retina_markers_df <-


motif_df$motif[match(retina_markers[[2]], motif_df$name)]
sObj <- readRDS("rObjects/shared_sObj_merge_with13PCW.rds")

```


## Annotate cells

The cells were annotated using the annotations on slide 155 with new AC definitions:

```{r}
anno <- read.csv("annotation/all_samples")

sObj$CellType <- anno$cell_type[match(sObj$peaks_snn_res.0.8, anno$cluster)]
DimPlot(sObj, group.by = "CellType", label = TRUE)
```


## All cells trajectory 


```{r}
start_Node = "Y_106"
```

Monocle 3 was used to calculate the pseudotime trajectory for all cells starting from `r start_Node`:

```{r}
cds_all <- as.cell_data_set(sObj)
cds_all <- cluster_cells(cds = cds_all, reduction_method = "UMAP")
cds_all <- learn_graph(cds_all, use_partition = FALSE, close_loop = FALSE)
cds_all <- order_cells(cds_all, reduction_method = "UMAP", root_pr_nodes = start_Node)
```

```{r}
plot_cells(
  cds = cds_all,
  show_trajectory_graph = TRUE,
  color_cells_by = "pseudotime", label_principal_points = TRUE
)
```

```{r}
sObj <- AddMetaData(
  object = sObj,
  metadata = cds_all@principal_graph_aux@listData$UMAP$pseudotime,
  col.name = "all_pseudotime"
)
```

The pseudotime estimated are shown on the original umap below:

```{r, fig.width=20}
p1 <- FeaturePlot(sObj, c("all_pseudotime"), pt.size = 0.1) & scale_color_viridis_c()  
p2 <- DimPlot(sObj, group.by = "CellType", label = TRUE)


p2 + p1 
```

## RPC T1 T2 T2 trajectory

The RPC, T1, T2, T3 cell types were selected and pseudotime analysis performed. 

```{r, eval = FALSE}

RPC_T1_T2_T3 <- sObj[,  sObj$CellType %in% c("early RPCs", "late RPCs", "T1", "T2", "T3")]


RPC_T1_T2_T3[,RPC_T1_T2_T3$nCount_peaks > 25000]


RPC_T1_T2_T3_new <- RunUMAP(object = RPC_T1_T2_T3,
                      reduction = 'lsi',
                      dims = 2:10,
                      n.neighbors = 30,
                      min.dist = 0.25,
                      n.epochs = 5000
                      )



saveRDS(RPC_T1_T2_T3_new, "rObjects/seurat_PT_RPC_T1_T2_T3.rds")



```

```{r}
start_Node = "Y_31"
RPC_T1_T2_T3_new <- readRDS("rObjects/seurat_PT_RPC_T1_T2_T3.rds")

cds <- as.cell_data_set(RPC_T1_T2_T3_new)
cds <- cluster_cells(cds = cds, reduction_method = "UMAP")
cds <- learn_graph(cds, use_partition = FALSE, close_loop = FALSE)
cds <- order_cells(cds, reduction_method = "UMAP", root_pr_nodes = start_Node)
```

```{r}

plot_cells(
  cds = cds,
  show_trajectory_graph = TRUE,
  color_cells_by = "CellType", label_principal_points = TRUE
)

```

Monocle 3 was used to calculate the pseudotime trajectory for all cells starting from `r start_Node`:

```{r}
cds <- order_cells(cds, reduction_method = "UMAP", root_pr_nodes = start_Node)



```



```{r}
plot_cells(
  cds = cds,
  show_trajectory_graph = TRUE,
  color_cells_by = "pseudotime", label_principal_points = TRUE
)

```
The pseudotime estimated are shown on the original umap below:

```{r}
sObj <- AddMetaData(
  object = sObj,
  metadata = cds@principal_graph_aux@listData$UMAP$pseudotime,
  col.name = "RPC_T1_T2_T3"
)

DefaultAssay(sObj) <-"chromvar"


```




```{r}

pseudotime <-sObj$RPC_T1_T2_T3[!is.na(sObj$RPC_T1_T2_T3)]
cor_row <- function(x){cor(pseudotime, x) }


addCorrelation <- function(cellType){
motif_mat <- sObj@assays$chromvar@data[motifs[[cellType]]$motif,!is.na(sObj$RPC_T1_T2_T3)]
motifs[[cellType]]$corr_pseudotime <- apply(motif_mat, 1, cor_row)}


motifs <- lapply(names(motifs), addCorrelation)





```


```{r, fig.width=20}
p1 <- FeaturePlot(sObj, c("RPC_T1_T2_T3"), pt.size = 0.1) & scale_color_viridis_c()  
p2 <- DimPlot(sObj, group.by = "CellType", label = TRUE)


p2 + p1 
```


```{r}
FeaturePlot(sObj, c("RPC_T1_T2_T3"), pt.size = 0.1) & scale_color_viridis_c()

FeaturePlot(sObj, c("nCount_peaks"), pt.size = 0.1) & scale_color_viridis_c()
DimPlot(sObj,group.by = "ID", pt.size = 0.1) 
```

