---
title: "10PCW_Retina"
output:
  html_document:
    df_print: paged
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```



```{r, libraries}

library(Seurat)
library(cowplot)
library(SCFunctionsV3)
library(dplyr)
library(tidyr)
library(stringr)
library(tidyverse)
library(DoubletFinder)
library(pheatmap)
library(DoubletFinder)

### Functions
howManyCells <- function(sObj){dim(sObj)[2]}


### Functions
preDoublets <- function(seuratObj){  
    seuratObj <- NormalizeData(seuratObj)
    
    seuratObj <- ScaleData(seuratObj)
    seuratObj <- FindVariableFeatures(seuratObj, 
                                      x.low.cutoff = 0.0125, 
                                      y.cutoff = 0.25, 
                                      do.plot=FALSE)
    seuratObj <- RunPCA(seuratObj, pc.genes = seuratObj@var.genes, pcs.print = 0)
    
    # set.seed(1234)
    # seuratObj <- RunTSNE(seuratObj, dims.use = 1:10, verbose=TRUE)
    return(seuratObj)
}

findDoublets <- function(seuratObj){
    
    ### calculate expected number of doublets
    nExp_poi  <- round(0.15*nrow(seuratObj@meta.data))
    
    ### predict doublets
    seuratObj <- doubletFinder_v3(seuratObj, 
                                  PCs = 1:10, 
                                  pN = 0.25, 
                                  pK = 0.01, 
                                  nExp = nExp_poi, 
                                  reuse.pANN = FALSE, 
                                  sct=FALSE)
    
    
    seuratObj@meta.data <- seuratObj@meta.data %>% 
        rename_at(vars(starts_with("DF.classifications")), 
                  funs(str_replace(., ".*", "DF.classifications"))) %>%
        rename_at(vars(starts_with("pANN")), 
                  funs(str_replace(., ".*", "pANN")))
    
    return(seuratObj) 
    
}

```

```{r}

sObj_List <- readRDS("rObjects/sObj_List.rds")
sObj_Filtered <- readRDS("rObjects/sObj_Filtered.rds")
sampleInfo <- readRDS("rObjects/sampleInfo.rds")
samples <- list.files("data") 
#integrated <- readRDS("rObjects/integrated.rds")
sObj <- readRDS("rObjects/clustered_sObj.rds")
clusterRes <- c(0.2, 0.4, 0.6, 0.8, 1)

```

Samples: 

1 x 10pcw retina: sequenced 2020_23, no id




```{r, eval = FALSE}
samples <- list.files("data")  
sampleInfo <- data.frame(sample = "10PCW_23_2020")
stage = "10PCW"
sampleInfo$stage <- rep(stage, length(samples))
sampleInfo$tissue <- rep("retina", length(samples))
sObj_List <- list()

for (i in 1:length(samples)){
  df <- Read10X(file.path("data", samples[i], "filtered_feature_bc_matrix"))
  sObj_List[[i]] <- CreateSeuratObject(counts = df, project = samples[i], min.cells = 10, min.features = 0)
  sObj_List[[i]][["percent.mt"]] <- PercentageFeatureSet(sObj_List[[i]], pattern = "^MT-")

  sObj_List[[i]]$sample <- sampleInfo$sample[i]
  sObj_List[[i]]$stage <- sampleInfo$stage[i]
  sObj_List[[i]]$tissue <- sampleInfo$tissue[i]
  
}


saveRDS(sObj_List, "rObjects/sObj_List.rds")

sampleInfo$beforeFiltering <- sapply(sObj_List, howManyCells)


```

```{r}
nCount_RNA_List <- list()
nFeature_RNA_List <- list()
mt_List <- list()

minCounts <- 1000
minFeatures <- 500
maxMit <- 15

for (i in 1:length(samples)){
  nCount_RNA_List[[i]] <- QCplotHist(sObj_List[[i]], "nCount_RNA", Xmin =  0  , Xmax = 80000,
    NumberOfBins = 20) + xlim(minCounts, 80000)
  
  nFeature_RNA_List[[i]] <- QCplotHist(sObj_List[[i]], "nFeature_RNA",  Xmax = NULL, Xmin = minCounts,
    NumberOfBins = 15)
  
  mt_List[[i]] <- FeatureScatter(object = sObj_List[[i]], 
                 feature1 = "nFeature_RNA", 
                 feature2 = "percent.mt") + 
       geom_hline(yintercept = maxMit)
}
```


## Number of Counts per cell

```{r}
plot_grid(plotlist = nCount_RNA_List, ncol = 1)
```

## Number of Gene per cell

```{r}
plot_grid(plotlist = nFeature_RNA_List, ncol = 1)
```

## Percentage mitochondrial reads

```{r, fig.height=15}
plot_grid(plotlist = mt_List, ncol = 1)
```


### Filtering

```{r, eval = FALSE}
sObj_Filtered <- list()
for (i in 1:length(samples)){
  filter <- sObj_List[[i]]$nCount_RNA > minCounts & 
    sObj_List[[i]]$nFeature_RNA > minFeatures & 
    sObj_List[[i]]$percent.mt < maxMit
  
  sObj_Filtered[[i]] <- sObj_List[[i]][ , filter]
  

}



### Find and remove doublets
for (i in 1:length(sObj_Filtered)){
    sObj_Filtered[[i]] <-  preDoublets(sObj_Filtered[[i]])
    sObj_Filtered[[i]] <- findDoublets(sObj_Filtered[[i]])
    
    cellFilter <- sObj_Filtered[[i]]$DF.classifications == "Singlet"
    sObj_Filtered[[i]] <- sObj_Filtered[[i]][, cellFilter]
    
    sObj_Filtered[[i]] <- NormalizeData(sObj_Filtered[[i]], verbose = FALSE)
    sObj_Filtered[[i]] <- FindVariableFeatures(sObj_Filtered[[i]], 
                                                 selection.method = "vst",
                                                 nfeatures = 2000, 
                                                 verbose = FALSE)
}



sampleInfo$afterFiltering <- sapply(sObj_Filtered, howManyCells)

saveRDS(sObj_Filtered, "rObjects/sObj_Filtered.rds")
saveRDS(sampleInfo, "rObjects/sampleInfo.rds")
```

The number of cells before and after filtering is shown below:

```{r}
sampleInfo
```




```{r, integration, eval = FALSE}

### Integration

#The datsets were combined in the usual way using the seurat anchor method.

anchors <- FindIntegrationAnchors(object.list = sObj_Filtered[1:3], dims = 1:10)

integrated <- IntegrateData(anchorset = anchors, dims = 1:10)

DefaultAssay(integrated) <- "integrated"

# Run the standard workflow for visualization and clustering
integrated <- ScaleData(integrated, verbose = TRUE)
integrated <- RunPCA(integrated, npcs = 30, verbose = FALSE)
integrated <- RunUMAP(integrated, reduction = "pca", dims = 1:10)









## Integration Results

#The samples did not overlap well after clustering analysis. This might be due to lots of changes at this time point or quality #control issues with the samples. 

DimPlot(integrated,  group.by = "sample")


Cells were clustered at a resolution of 0.2, 0.4, 0.6, 0.8, 1, resulting in between 12 and 24 clusters. Markers were then identified for each resolution. 

p_list <- list()
for (cluster in clusterRes){
p_list[[as.character(cluster)]] <- DimPlot(integrated, group.by = paste0("integrated_snn_res.", cluster), label = TRUE) + NoLegend() + ggtitle(paste0("Resolution ", cluster))
}

plot_grid(plotlist = p_list, ncol = 1)





DefaultAssay(integrated) <- "RNA"

marker_list <- list()


for (cluster in clusterRes){
integrated <- SetIdent(integrated, value = paste0("integrated_snn_res.", cluster))
marker_list[[as.character(cluster)]] <- FindAllMarkers(integrated, 
                                                       only.pos = TRUE, 
                                                       min.pct = 0.25, 
                                                       logfc.threshold = 0.25)

marker_list[[as.character(cluster)]] <- marker_list[[as.character(cluster)]] %>% 
  group_by(cluster) %>% 
  arrange(cluster, desc(avg_logFC))

write.csv(marker_list[[as.character(cluster)]], 
          paste0("markers/markers_res_", cluster, ".csv"))}

```



### Clustering 

A clustering analysis was performed on the single sample

```{r, eval = FALSE}

sObj <- sObj_Filtered[[1]]
# Run the standard workflow for visualization and clustering
sObj <- ScaleData(sObj, verbose = TRUE)
sObj <- RunPCA(sObj, npcs = 30, verbose = FALSE)
sObj <- RunUMAP(sObj, reduction = "pca", dims = 1:10)

sObj <- FindNeighbors(sObj, dims = 1:10, verbose = FALSE)

sObj <- FindClusters(sObj, resolution = clusterRes, verbose = FALSE)
saveRDS(sObj, "rObjects/clustered_sObj.rds")

```


Cells were clustered at a resolution of 0.2, 0.4, 0.6, 0.8, 1, resulting in between 10 and 21 clusters. Markers were then identified for each resolution. 
```{r, fig.height=20}
p_list <- list()
for (cluster in clusterRes){
p_list[[as.character(cluster)]] <- DimPlot(sObj, group.by = paste0("RNA_snn_res.", cluster), label = TRUE) + NoLegend() + ggtitle(paste0("Resolution ", cluster))
}

plot_grid(plotlist = p_list, ncol = 1)

```


```{r, eval = FALSE}
DefaultAssay(sObj) <- "RNA"

marker_list <- list()


for (cluster in clusterRes){
sObj <- SetIdent(sObj, value = paste0("RNA_snn_res.", cluster))
marker_list[[as.character(cluster)]] <- FindAllMarkers(sObj, 
                                                       only.pos = TRUE, 
                                                       min.pct = 0.25, 
                                                       logfc.threshold = 0.25)

marker_list[[as.character(cluster)]] <- marker_list[[as.character(cluster)]] %>% 
  group_by(cluster) %>% 
  arrange(cluster, desc(avg_logFC))

write.csv(marker_list[[as.character(cluster)]], 
          paste0("markers/markers_res_", cluster, ".csv"))}

```

